---
title: "reading_data_from_the_web"
author: "Purnima Sharma"
date: "10/16/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

library(rvest)
library(httr)  # contains collection of tools for constructing http requests.
```

##Scrape table
Load data from the web.

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)   # read html from that link

drug_use_html
```

Extracting tables from HTML.

```{r}
drug_use_html %>%
  html_nodes(css = "table")   # this code extracted all tables!
```

Extracting first table only.

```{r}
table_marj = 
  drug_use_html %>% 
  html_nodes(css = "table") %>% 
  first() %>%
  html_table() 
```

Remove first row from table(with same NOTe in every column), and converting the table to a df(tibble).

```{r}
table_marj = 
  drug_use_html %>% 
  html_nodes(css = "table") %>% 
  first() %>% 
  html_table() %>%
  slice(-1) %>% 
  as_tibble()

table_marj
```

------------------------------------------------------------------------------
Learning assessment 1: Create a data frame that contains the cost of living table for New York from the webp age.

```{r Load data from the web}
url = "https://www.bestplaces.net/cost_of_living/city/new_york/new_york"
cost_of_living_ny_html = read_html(url)

cost_of_living_ny_html 
```

```{r Extract table from html}
cost_of_living_ny_html %>%
  html_nodes(css = "table")  # all tables extracted!
```

Extracting first table only.

```{r}
table_ny = 
  cost_of_living_ny_html %>% 
  html_nodes(css = "table") %>% 
  first() %>%
  html_table() 
```

Converting to tibble (df).

```{r}
table_ny = 
  cost_of_living_ny_html %>% 
  html_nodes(css = "table") %>% 
  first() %>%
  html_table() %>%
  as_tibble()

table_ny
```
OR, code for assessment 1:
```{r}
nyc_cost = 
  read_html("https://www.bestplaces.net/cost_of_living/city/new_york/new_york") %>%
  html_nodes(css = "table") %>%
  .[[1]] %>%
  html_table(header = TRUE)
```
End assessment 1 -------------------------------------------------------------

## CSS selectors

Star Wars example: Getting data about the Star Wars Movies from the IMDB page.

```{r load data from the web}
swm_html = 
  read_html("https://www.imdb.com/list/ls070150896/")
```

Not a handy table, so using CSS selector to isolate elements needed.
For each element, use the CSS selector in html_nodes() to extract the relevant HTML code, and convert it to text. Then combine these into a data frame.

```{r}
title_vec = 
  swm_html %>%
  html_nodes(".lister-item-header a") %>%
  html_text()

gross_rev_vec = 
  swm_html %>%
  html_nodes(".text-small:nth-child(7) span:nth-child(5)") %>%
  html_text()

runtime_vec = 
  swm_html %>%
  html_nodes(".runtime") %>%
  html_text()

swm_df = 
  tibble(
    title = title_vec,
    rev = gross_rev_vec,
    runtime = runtime_vec)
```

----------------------------------------------------------------------------
Learning Assessment 2: extracting titles of the reviews for "Napoleon Dynamite" from the given website.

```{r load data from the web}
nd_html = 
  read_html("https://www.amazon.com/product-reviews/B00005JNBQ/ref=cm_cr_arp_d_viewopt_rvwer?ie=UTF8&reviewerType=avp_only_reviews&sortBy=recent&pageNumber=1")
```

Extracting titles, stars and text of reviews

```{r}
review_titles = 
  nd_html %>%
  html_nodes(".a-text-bold span") %>%
  html_text()

review_stars = 
  nd_html %>%
  html_nodes(".review-rating") %>%
  html_text()

review_text = 
  nd_html %>%
  html_nodes(".review-text-content span") %>%
  html_text()

nd_df = 
  tibble(
    title = review_titles,
                           
    text = review_text 
  )  ## stars rating not working in the table, it has 12obs and these two have 10obs.
```
End Learning assessment 2 ----------------------------------------------------


Using API (Application Programming Interface), instead of reading through a website as is:

Example 1
a dataset for annual water consumption in NYC. import this as a CSV and parse it.

```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") %>% 
  content("parsed")
```

Importing this data as a JSON file (A JSON file is a file that stores simple data structures and objects in JavaScript Object Notation (JSON) format, which is a standard data interchange format. It is primarily used for transmitting data between a web application and a server.)
 Package used: "JSONLITE"

```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.json") %>% 
  content("text") %>%
  jsonlite::fromJSON() %>%
  as_tibble()
```

Example 2
Reading data from Data.gov:  selecting data coming from BRFSS. This is importable via the API as a CSV

```{r}
brfss_smart2010 = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",  #gets 5000 rows vs default 1000.
      query = list("$limit" = 5000)) %>% 
  content("parsed")
```


Example 3
Reading pokemon API

```{r}
poke = 
  GET("http://pokeapi.co/api/v2/pokemon/1") %>%
  content()

poke$name    #output"bulbasaur"
poke$height   # output : 7
poke$abilities

# To build a Pokemon dataset for analysis, you’d need to distill the data returned from the API into a useful format; iterate across all pokemon; and combine the results.
```





