---
title: "strings_and_factors"
author: "Purnima Sharma"
date: "10/17/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)
```

Global settings for themes (for figures).
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,          #aspect ratio: height over width
  out.width = "90%"      # figure doesn't go all the way out on full(100%) page
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", # to override ggplot default colors.
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## strings as vectors.

```{r}
string_vec = c("my", "name", "is", "purnima")
str_detect(string_vec, "purnima")           #using detect to find match

str_replace(string_vec, "purnima", "Purnima") #replace matched value.
```

Designate matches beginning or end of a line.

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")   # detect match beginning of line.

str_detect(string_vec, "i think$")  #end of line
```

LIst of characters as a match.

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush") # match lower case "ush" with any B or b
```

Provide a range of letters or numbers that count as a match.

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]") #matching with beginning with these#s
```

The character . matches anything.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")  #match 7 "something" eleven. 3rd sentence didn't have "something"
```

Special characters, [ and ], ( and ), and . are detected using \ .But \ is also special, so use \ twice!

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")      
```


## factors

EXAMPLLE 1: NSDUH ( National Survey on Drug Use and Health) data.

```{r reading data from the web}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_nodes(css = "table") %>% 
  first() %>%
  html_table() %>%
  slice(-1) %>%
  as_tibble()
```

Tidy data.

```{r}
data_marj = 
  table_marj %>%
  select(-contains("P Value")) %>%    # remove p-values
  pivot_longer( 
    -State,                          
    names_to = "age_year",        # fixing the column headers, except state column.   
    values_to = "percent") %>%
  separate(age_year, into = c("age", "year"), sep = "\\(") %>%
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),  #removing letters from table's percent values.
    percent = as.numeric(percent)) %>%
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

Focusing on 12 - 17 age group.

```{r}
data_marj %>%
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>%    # state as factor, and reorder by median percent values. 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
End Example 1 --------------------------------------


EXAMPLE 2: NYC Restuarant Inspections data.----------

Import data

```{r}
data("rest_inspec")

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  spread(key = grade, value = n)
```

Remove inspections with scores other than A, B, or C, and remove the restaurants with missing boro information, and clean boro names.

```{r}
rest_inspec =
  rest_inspec %>%
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro))
```

Focusing on pizza places and reexamining grades by boro.

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  spread(key = grade, value = n)
```

Above code did not give all the pizza places, since "detect" is case sensitive. Recoding by specifying detect:

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%  # accept all forms of writing "pizza"
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  spread(key = grade, value = n)
```

Visualizing above results.

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  ggplot(aes(x = boro, fill = grade)) +    #fill each bar colors by grades.
  geom_bar() 
```

Rearranging bars according to number of pizza places(frequency).

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(boro = fct_infreq(boro)) %>%       # arranging boro-bars by frequency, using fct_infreq
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```

Renaming a boro.

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Brooklyn", "HipsterHQ")) %>%  #rename Brooklyn
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```
Renaming like this converted boros back to string and arranged alphabetically on graph.

Code to rename while retaining factor levels:

```{r}
rest_inspec %>% 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) %>% # another way to ignore upper/lower case (other than specifying as above)
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "HipsterHQ" = "Brooklyn")) %>%  # recode instead of replace to rename it.
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```
End example 2 ----------------------------------


Example 3: Weather data ------------------------

```{r load data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Reorder name according to tmax values in each name using fct_reorder:

```{r}
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
End example 3 -----------------------------


Example 4: Pulse data (processing using stringr and forcats packages)--------

```{r}           
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = str_replace(visit, "bl", "00m"),
    visit = fct_relevel(visit, str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

print(pulse_data, n = 12)
```
End example 4 ---------------------------------


Example 5: Air bnb data (using "factors" to order neighborhoods by price) ----

```{r}
data("nyc_airbnb")

nyc_airbnb %>%
  filter(neighbourhood_group == "Manhattan") %>% 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, price)) %>% 
  ggplot(aes(x = neighbourhood, y = price)) +
  geom_boxplot() +
  coord_flip() + 
  ylim(0, 1000)
```

